---
title: | 
  **Supplemental Materials** for "Statements issued by academic medical
  institutions after George Floyd’s killing by police and subsequent 
  unrest in the United States: cross-sectional study"
author: Mathew V Kiang and Alexander C Tsai
output: 
  html_document:
    code_folding: 'hide'
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r imports}
library(tidyverse)
library(here)
library(DT)
library(knitr)
```

```{r}
## Read in processed data
analytic_df <- readRDS(here("data_private", "social_mission_analytic_df.RDS"))
word_df <- readRDS(here("data_private", "social_mission_tokenized_words.RDS"))
sentence_df <- readRDS(here("data_private", "social_mission_tokenized_sentences.RDS"))
ngram2_df <- readRDS(here("data_private", "social_mission_tokenized_bigrams.RDS"))
ngram3_df <- readRDS(here("data_private", "social_mission_tokenized_trigrams.RDS"))
summary_df <- readRDS(here("data_private", "social_mission_keyword_summary_df.RDS"))
```

## List institutions included in the data set

A total of `r n_distinct(analytic_df$institution)` statements are included in the data set. The institutions, listed in alphabetical order, are:

```{r}
analytic_df %>%
  arrange(institution) %>%
  transmute(institution,
            institution_star =
              paste0(
                institution,
                ifelse(public == 0, "*", ""),
                ifelse(!grepl("[Dd]{1}ean", position), "^", "")
              ),
            mullan_rank, 
            nih_rank,
            usnwr_rank) %>%
  select(-institution_star) %>%
  knitr::kable(
    format = "html",
    digits = 0,
    col.names = c("Institution", "Social Mission Rank (Mullan)", "NIH Funding Rank (2018)", "USN&WR Rank (2021)")
  )
```

Of the `r n_distinct(analytic_df$institution)` statements, `r sum(analytic_df$public == 0)` (`r sprintf("%0.1f%%", round(sum(analytic_df$public == 0) / n_distinct(analytic_df$institution) * 100, 1))`) were not easily accessible on the institution's website or official social media accounts and `r sum(!grepl("[Dd]{1}ean", analytic_df$position))` (`r sprintf("%0.1f%%", round(sum(!grepl("[Dd]{1}ean", analytic_df$position)) / n_distinct(analytic_df$institution) * 100, 1))`) were not from a medical school dean. 

## Dates of statements

All statements occurred between `r min(analytic_df$date, na.rm = TRUE)` and `r max(analytic_df$date, na.rm = TRUE)` with `r x <- sum(is.na(analytic_df$date)); x` undated `r ifelse(x > 1, "statements", "statement")`.

```{r}
analytic_df %>%
  count(date)
```

The median and IQR of dates is:

```{r}
sort(analytic_df$date)[c(round(.5 * NROW(analytic_df)), 
                         round(.25 * NROW(analytic_df)),
                         round(.75 * NROW(analytic_df)))]
```

## Descriptive statistics of statements

### Total number of unique words (excluding stop words)

```{r}
word_df %>% 
  filter(keep_word) %>% 
  pull(word) %>% 
  n_distinct()
```

### Total number of words (including repeats, excluding stop words)

```{r}
word_df %>% 
  filter(keep_word) %>% 
  NROW()
```

### Total number of sentences

```{r}
sentence_df %>% 
  NROW()
```

Note that counting sentences is a rough approximation because things like "Mr." will count as their own sentence.

### Words by institution (including repeats and stop words)

```{r}
x <- word_df %>% 
    count(random_id, sort = TRUE)

ggplot(x, aes(x = random_id, y = n)) + 
    geom_col() + 
    theme_light()
```

```{r}
summary(x$n)
```

### Words by institution (including repeats, excluding stop words)

```{r}
x <- word_df %>% 
    filter(keep_word) %>% 
    count(random_id, sort = TRUE)

ggplot(x, aes(x = random_id, y = n)) + 
    geom_col() + 
    theme_light()
```

```{r}
summary(x$n)
```

## Most common words and n-grams

### 100 most common words and word stems

Here, we tabulate the frequency of each `word_stem` and `word`. The `n_word_stem` and `n_word` columns show the frequency of stems and words, respectively. The `n_inst_*` columns show the number of unique institutes that used each word. Sorting by `n_inst_word` will result in the words that were used by the highest number of different institutions. 

```{r}
x <- word_df %>%
    filter(keep_word) %>%
    group_by(word_stem) %>%
    mutate(n_word_stem = n(),
           n_inst_word_stem = n_distinct(random_id)) %>%
    group_by(word) %>%
    mutate(n_word = n(), n_inst_word = n_distinct(random_id)) %>%
    select(word_stem,
           word,
           n_word_stem,
           n_inst_word_stem,
           n_word,
           n_inst_word) %>%
    distinct() %>% 
    ungroup() %>% 
    mutate(stem_rank = min_rank(desc(n_word_stem))) %>% 
    filter(stem_rank <= 100) %>% 
    arrange(desc(n_word_stem))

x %>% 
    datatable(rownames = FALSE)
```

### 100 most common 2-gram phases

Below, we show the top 100 most commonly used 2-gram phrases in both the stemmed and unstemmed format. Similar to above, we show the counts (`n_ngram_stem` for the stemmed version and `n_ngram` for the unstemmed version) as well as the number of unique institutions that used the phrase (e.g., `n_inst_ngram_stem` for the stemmed version). 

```{r}
x <- ngram2_df %>% 
    group_by(ngram2_stem) %>%
    mutate(n_ngram_stem = n(),
           n_inst_ngram_stem = n_distinct(random_id)) %>%
    group_by(ngram2) %>%
    mutate(n_ngram = n(), 
           n_inst_ngram = n_distinct(random_id)) %>%
    select(ngram2_stem,
           ngram2,
           n_ngram_stem,
           n_inst_ngram_stem,
           n_ngram,
           n_inst_ngram) %>%
    distinct() %>% 
    ungroup() %>% 
    mutate(stem_rank = min_rank(desc(n_ngram_stem))) %>% 
    filter(stem_rank <= 100) %>% 
    arrange(desc(n_ngram_stem))

x %>% 
    datatable(rownames = FALSE)
```

### Most common 3-gram phrases

Like above but for three word phrases

I filter on the phrase being said more than two times. 

```{r}
x <- ngram3_df %>% 
    group_by(ngram3_stem) %>%
    mutate(n_ngram_stem = n(),
           n_inst_ngram_stem = n_distinct(random_id)) %>%
    group_by(ngram3) %>%
    mutate(n_ngram = n(), 
           n_inst_ngram = n_distinct(random_id)) %>%
    select(ngram3_stem,
           ngram3,
           n_ngram_stem,
           n_inst_ngram_stem,
           n_ngram,
           n_inst_ngram) %>%
    distinct() %>% 
    ungroup() %>% 
    filter(n_ngram_stem > 1) %>% 
    arrange(desc(n_ngram_stem))

x %>% 
    datatable(rownames = FALSE)
```
